---
title: "Multi omics real data"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 100, 
                      echo=F, 
                      warning=F, message=F, eval = TRUE,
                      fig.show=TRUE, fig.width= 5,fig.height= 5, out.width = '50%', fig.ncol = 1, fig.path= 'Figures/')

setwd("./plsko_writing/paper codes/multiomics_semi_casestudy")

library(edgeR)
library(limma)
library(tidyverse)
library(mixOmics)
library(knockoff)
require(glmnet)
require(graphics)
library(reshape2)
library(rARPACK)
library(doParallel)

source.path <- "./plsko_writing/paper codes/RSources"
source(paste0(source.path, "data_sim1.R"))
source(paste0(source.path, "plsko_plsonly.R"))
source(paste0(source.path, 'KO_perf.R'))
source(paste0(source.path, 'KOBT/create.pc.knockoff.R'))
source(paste0(source.path, 'hdko.R'))
source(paste0(source.path, 'mvr_julia.R'))
source(paste0(source.path, 'ako_run.R'))
source(paste0(source.path, 'ako_with_ko_run.R'))


julia <- julia_setup()

cl <- makeCluster(11) 
registerDoParallel(cl)
```

Data were downloaded from: https://github.com/ivanam5/Multiomics_ Preeclampsia
Please refer to the original paper for more information: Marić I, Contrepois K, Moufarrej MN et al. Early prediction and longitudinal modeling of preeclampsia from multiomics. PATTER 2022;3, DOI: 10.1016/j.patter.2022.100655.

```{r loading data}
load("data/raw_data.rda")
sample.info <- read.csv("./data/proteome_edited.csv")

sample.id <- colnames(metabolome_urine_data_raw)[1:172]
sub.id <- stringr::str_sub(sample.id, 2, 6)

sample.group.time1 <- proteome_data_raw[,1:3] %>% arrange(Subject_ID) %>% filter(TimePoint=="1st Tri")

sample.time1 <- data.frame(sample.id, Subject_ID = sub.id) %>%  filter(!duplicated(Subject_ID)) %>% filter(.$Subject_ID %in% sample.group.time1$Subject_ID) %>% arrange(Subject_ID) %>% merge(sample.group.time1)

#batch information
batch1 <- read.csv("patientsalldata6.csv")
batch2 <- read.csv("patients16.csv")

```
#Semi-synthetic data
## proteomics
```{r preprocess}
prot.time1 <- proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",-c(1:3)] %>% log2()
rownames(prot.time1) <- proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri","Subject_ID"]

# batcheffect removal
batch.prot <- as.numeric(rownames(prot.time1)%in% batch1$x)
prot.time1.batrm <- limma::removeBatchEffect(t(prot.time1), batch = batch.prot)
prot.time1.batrm <- t(prot.time1.batrm)
```


```{r ko semi sim}
# p = 50, 100, 150
cl <- makeCluster(11) 
registerDoParallel(cl)

r_emp_X <- r_criterion(prot.time1, rmax = 10)

# p_x = 100, 150, 200, p_s = 25
it = 50
test <- foreach(i = rep(1:it,3), px = rep(seq(40, 120, 40), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             res <- real_sim(prot.time1.batrm, p_x = px, p_s = 15, ko.diagnosis = F, seed = i, call.var = "p_x", plsko.ncomp = 5, plsko.threshold.q = 0.8, q = 0.1)
                           }
test_df <- do.call("rbind",  test )
saveRDS(test_df, "./prot_real.rds")

```

```{r plot}
library(tidyverse)
library(ggpubr)
library(viridis)
library(scales)
library(mixOmics)
test_df <- readRDS("./prot_real.rds")
test_df$P <- as.factor(test_df$var)
test_df$ko.method[test_df$ko.method == "plsonly"] <- "PLSKO"
test_df$ko.method <- factor(test_df$ko.method)

col_scip <- method.colour[levels(test_df$ko.method)]


test_p_fdr.plot <- ggplot(mapping = aes(x = P, y = fdp), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = scip_fullnames, values = col_scip)+
  scale_color_manual(labels = scip_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")+
  ylab("FDP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test_p_fdr.plot

test_p_pow.plot <- ggplot(mapping = aes(x = P, y = tpp), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = scip_fullnames, values = col_scip)+
  scale_color_manual(labels = scip_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")+
  ylab("TPP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test_p_pow.plot

prot.semi <- ggarrange(test_p_fdr.plot, test_p_pow.plot, nrow = 1, ncol = 2, common.legend = T)

# plus 
test_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test_p_fdr_plus.plot

test_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test_p_pow_plus.plot

# run time
test_p_time.plot <- ggplot(mapping = aes(x = P, y = run.time), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test_p_time.plot
```
### mvr
```{r mvr test, eval=FALSE}

it = 50
set.seed(1)

test.mvr <- foreach(i = rep(1:it,3), px = rep(seq(40, 120, 40), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab","JuliaCall")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             julia_library("Knockoffs")
                             res <- real_mvr(prot.time1.batrm, p_x = px, p_s = 25, ko.diagnosis = F, seed = i, call.var = "p_x", q = 0.1)
                           }
real_mvr_df <- do.call("rbind",  test.mvr )
saveRDS(real_mvr_df, "./prot_mvr.rds")
```

```{r mvr plot}
real_all_mvr_df <- readRDS("./prot_mvr.rds")
real_all_mvr_df$P <- factor(real_all_mvr_df$var, levels = c("40","80","120"))
real_all_mvr_df$ko.method[real_all_mvr_df$ko.method=="sdp_JS"] <- "JS"
real_all_mvr_df$ko.method <- factor(real_all_mvr_df$ko.method)
col_scip <- method.colour[levels(real_all_mvr_df$ko.method)]

real_all_mvr_df$ko.method <- factor(real_all_mvr_df$ko.method)

test.mvr_p_fdr.plot <- ggplot(mapping = aes(x = P, y = fdp), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = mvr_fullnames, values = col_scip)+
  scale_color_manual(labels = mvr_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")+
  ylab("FDP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test.mvr_p_fdr.plot

test.mvr_p_pow.plot <- ggplot(mapping = aes(x = P, y = tpp), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = mvr_fullnames, values = col_scip)+
  scale_color_manual(labels = mvr_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")+
  ylab("TPP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test.mvr_p_pow.plot

prot.semi.mvr <-  ggarrange(test.mvr_p_fdr.plot, test.mvr_p_pow.plot, nrow = 1, ncol = 2, common.legend = T)

ggarrange(prot.semi, prot.semi.mvr, nrow = 2, ncol = 1, common.legend = F)

# plus 
test.mvr_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test.mvr_p_fdr_plus.plot

test.mvr_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr_p_pow_plus.plot

# run time
test.mvr_p_time.plot <- ggplot(mapping = aes(x = p_var, y = run.time), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr_p_time.plot
```
### mvr norm
```{r mvr norm}

test.mvr.norm <- foreach(i = rep(1:it,3), px = rep(seq(40, 120, 40), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab","JuliaCall")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             julia_library("Knockoffs")
                             res <- real_mvr_norm(prot.time1.batrm, p_x = px, p_s = 15, ko.diagnosis = F, seed = i, call.var = "p_x", q = 0.1)
                           }
real_mvr.norm_df <- do.call("rbind",  test.mvr.norm )
saveRDS(real_mvr.norm_df, "./real_mvr_norm.rds")
```

```{r mvr plot norm}
real_all_mvr.norm_df <- readRDS("./real_mvr_norm.rds")
real_all_mvr.norm_df$p_var <- factor(real_all_mvr.norm_df$var, levels = c("40","80","120"))
real_all_mvr.norm_df$ko.method <- factor(real_all_mvr.norm_df$ko.method)

test.mvr.norm_p_fdr.plot <- ggplot(mapping = aes(x = p_var, y = fdp), data = real_all_mvr.norm_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test.mvr.norm_p_fdr.plot

test.mvr.norm_p_pow.plot <- ggplot(mapping = aes(x = p_var, y = tpp), data = real_all_mvr.norm_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr.norm_p_pow.plot

# plus 
test.mvr.norm_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = real_all_mvr.norm_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test.mvr.norm_p_fdr_plus.plot

test.mvr.norm_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = real_all_mvr.norm_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr.norm_p_pow_plus.plot

# run time
test.mvr.norm_p_time.plot <- ggplot(mapping = aes(x = p_var, y = run.time), data = real_all_mvr.norm_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr.norm_p_time.plot
```
## metabolome (urine)
```{r metabolome urine data process }
metabo <- metabolome_urine_data_raw[,1:172] %>% log2()
metabo.sample <- colnames(metabo)
metabo.sub <- stringr::str_sub(metabo.sample, 2,6)

time1.sample <- (!duplicated(metabo.sub))&(metabo.sub %in% as.character(sample.group.time1$Subject_ID))

metabo.time1 <- metabo[,time1.sample] 
#rownames(metabo.time1) <- metabolome_urine_data_raw$Compounds_ID
metabo.time1 <- t(metabo.time1)
rownames(metabo.time1) <- metabo.sub[time1.sample]

# batcheffect removal
plotIndiv(pca(metabo.time1), group = as.factor(row.names(metabo.time1)%in%batch1$x))
batch.prot <- as.numeric(rownames(metabo.time1)%in% batch1$x)
metabo.time1.batrm <- limma::removeBatchEffect(t(metabo.time1), batch = batch.prot)
metabo.time1.batrm <- t(metabo.time1.batrm)
plotIndiv(pca(metabo.time1.batrm), group = as.factor(row.names(metabo.time1)%in%batch1$x))

```

```{r ko semi sim metabolome}
# p = 50, 100, 150
cl <- makeCluster(11) 
registerDoParallel(cl)

r_emp_X <- r_criterion(metabo.time1, rmax = 10)

# p_x = 100, 150, 200, p_s = 25
it = 50
test <- foreach(i = rep(1:it,3), px = rep(seq(40, 120, 40), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             res <- real_sim(metabo.time1, p_x = px, p_s = 15, ko.diagnosis = F, seed = i, call.var = "p_x", plsko.ncomp = 5, plsko.threshold.q = 0.8, q = 0.1)
                           }
test_df <- do.call("rbind",  test )
saveRDS(test_df, "./metabolome_real.rds")

```

```{r meta plot}
test_df <- readRDS("./metabolome_real.rds")
test_df$P <- as.factor(test_df$var)
test_df$ko.method[test_df$ko.method == "plsonly"] <- "PLSKO"
test_df$ko.method <- factor(test_df$ko.method)

col_scip <- method.colour[levels(test_df$ko.method)]


test_p_fdr.plot <- ggplot(mapping = aes(x = P, y = fdp), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = scip_fullnames, values = col_scip)+
  scale_color_manual(labels = scip_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")+
  ylab("FDP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test_p_fdr.plot

test_p_pow.plot <- ggplot(mapping = aes(x = P, y = tpp), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = scip_fullnames, values = col_scip)+
  scale_color_manual(labels = scip_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")+
  ylab("TPP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test_p_pow.plot

metabol.semi <- ggarrange(test_p_fdr.plot, test_p_pow.plot, nrow = 1, ncol = 2, common.legend = T)

# plus 
test_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test_p_fdr_plus.plot

test_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test_p_pow_plus.plot

# run time
test_p_time.plot <- ggplot(mapping = aes(x = p_var, y = run.time), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test_p_time.plot
```

### mvr

```{r mvr test metabolome, eval=FALSE}

it = 50
set.seed(1)

test.mvr <- foreach(i = rep(1:it,3), px = rep(seq(40, 120, 40), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab","JuliaCall")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             julia_library("Knockoffs")
                             res <- real_mvr(metabo.time1, p_x = px, p_s = 25, ko.diagnosis = F, seed = i, call.var = "p_x", q = 0.1)
                           }
real_mvr_df <- do.call("rbind",  test.mvr )
saveRDS(real_mvr_df, "./metabolome_mvr.rds")
```

```{r mvr plot metabolome}
real_all_mvr_df <- readRDS("./metabolome_mvr.rds")
real_all_mvr_df$P <- factor(real_all_mvr_df$var, levels = c("40","80","120"))
real_all_mvr_df$ko.method[real_all_mvr_df$ko.method=="sdp_JS"] <- "JS"
real_all_mvr_df$ko.method <- factor(real_all_mvr_df$ko.method)
col_scip <- method.colour[levels(real_all_mvr_df$ko.method)]

real_all_mvr_df$ko.method <- factor(real_all_mvr_df$ko.method)

test.mvr_p_fdr.plot <- ggplot(mapping = aes(x = P, y = fdp), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = mvr_fullnames, values = col_scip)+
  scale_color_manual(labels = mvr_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")+
  ylab("FDP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test.mvr_p_fdr.plot

test.mvr_p_pow.plot <- ggplot(mapping = aes(x = P, y = tpp), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = mvr_fullnames, values = col_scip)+
  scale_color_manual(labels = mvr_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")+
  ylab("TPP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test.mvr_p_pow.plot

metabol.semi.mvr <-  ggarrange(test.mvr_p_fdr.plot, test.mvr_p_pow.plot, nrow = 1, ncol = 2, common.legend = T)

ggarrange(metabol.semi, metabol.semi.mvr, nrow = 2, ncol = 1, common.legend = F)

# plus 
test.mvr_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test.mvr_p_fdr_plus.plot

test.mvr_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr_p_pow_plus.plot

# run time
test.mvr_p_time.plot <- ggplot(mapping = aes(x = p_var, y = run.time), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr_p_time.plot
```

## microbiome
```{r microbiome preprocess}

# filtering variables with excessive 0s
microbiome_counts <- microbiome_data_raw[,-c(1:12)]

# examine almost identical variables and only keep one for each highly-correlated cluster
remove_highly_correlated <- function(data, thres = 0.95) {
  # Compute the correlation matrix between variables
  cor_matrix <- cor(data)
  
  # Create a logical matrix to mark which columns need to be removed
  remove_cols <- logical(ncol(cor_matrix))
  
  for (i in 1:(ncol(cor_matrix) - 1)) {
    for (j in (i + 1):ncol(cor_matrix)) {
      # Check if the correlation is above the threshold
      if (abs(cor_matrix[i, j]) > thres) {
        # If the correlation is above the threshold (0.95), mark the variable to be removed
        if (!remove_cols[j]) {
          remove_cols[j] <- TRUE
        }
      }
    }
  }
  
  # Remove variables with high correlation from the data frame
  data_subset <- data[, !remove_cols]
  
  return(data_subset)
}

# remove exactly same columns
mb.counts <- remove_highly_correlated(microbiome_counts, thres = 1-10e-6) #820 var left
#test
# cor.mb <- cor(mb.counts)
# diag(cor.mb) <- 0
# heatmap(cor.mb,scale = "none")
# range(unlist(cor.mb))

#index.keep <- which(colSums(mb.counts)*100/(sum(colSums(mb.counts))) > 0.01)

# remove low abundant OTUs in time 1 
index.keep <- which(colSums(mb.counts[microbiome_data_raw$timepoint==1,])*100/(sum(colSums(mb.counts[microbiome_data_raw$timepoint==1,]))) > 0.01)


microbiome.keep <- mb.counts[, index.keep] #162 OTUs left

# add offset
microbiome.keep <- microbiome.keep+0.01

# CLR Transformation
mb.clr <- logratio.transfo(microbiome.keep, logratio = 'CLR')
class(mb.clr) <- 'matrix' 

# filter time 1 samples
mb.time1 <- mb.clr[microbiome_data_raw$timepoint==1,]
rownames(mb.time1) <- microbiome_data_raw$individual[microbiome_data_raw$timepoint==1]


#batch effect removal
plotIndiv(pca(mb.time1), group = as.factor(row.names(mb.time1)%in%batch1$x))
batch.prot <- as.numeric(rownames(mb.time1)%in% batch1$x)
mb.time1.batrm <- limma::removeBatchEffect(t(mb.time1), batch = batch.prot)
mb.time1.batrm <- t(mb.time1.batrm)
plotIndiv(pca(mb.time1.batrm), group = as.factor(row.names(mb.time1.batrm)%in%batch1$x))

```

```{r taxonomic assignment, eval = F}
library(dada2)

mb.taxa <- assignTaxonomy(colnames(mb.time1), "data/silva_nr99_v138.1_wSpecies_train_set.fa") # please find in https://zenodo.org/records/4587955
saveRDS(mb.taxa, "microbiome_taxa.rds")

mb.taxa <- readRDS("microbiome_taxa.rds")
```


```{r process highliy corrlated vars}
# remove low variable OTUs in time 1 (they might have high abundance in later time points but not time 1)
index.keep.time1 <- which(apply(mb.time1.batrm, 2, sd)!=0)

microbiome.keep.time1 <- mb.time1.batrm[, index.keep.time1] #162 OTUs left

#task 8 failed - "system is computationally singular: reciprocal condition number = 1.28389e-32
cor.mb.time1 <- cor(microbiome.keep.time1)
diag(cor.mb.time1) <- 0
hist(unlist(cor.mb.time1), breaks = 50)
range(unlist(cor.mb.time1)) # there still are variables which are almost the same (max cor= 0.9999972)

# remove highly-correlated variable again
mb.time1.rm <- remove_highly_correlated(microbiome.keep.time1, thres = 0.95)

#144 variable left

```

```{r ko semi sim mb}
# p = 50, 100, 150
cl <- makeCluster(11) 
registerDoParallel(cl)

r_emp_X <- r_criterion(mb.time1.rm, rmax = 10)

# p_x = 50, 75, 98, p_s = 25
it = 50
test <- foreach(i = rep(1:it,3), px = rep(c(50, 100, ncol(mb.time1.rm)), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             res <- real_sim(mb.time1.rm, p_x = px, p_s = 15, ko.diagnosis = F, seed = i, call.var = "p_x", plsko.ncomp = 5, plsko.threshold.q = 0.8, q = 0.1, rmax = 10)
                           }
test_df <- do.call("rbind",  test )
saveRDS(test_df, "./mb_real_144.rds")

```

```{r plot mb}
library(tidyverse)
library(ggpubr)
library(viridis)
library(scales)
library(mixOmics)
test_df <- readRDS("./mb_real_144.rds")
test_df$P <- as.factor(test_df$var)
test_df$ko.method[test_df$ko.method == "plsonly"] <- "PLSKO"
test_df$ko.method <- factor(test_df$ko.method)

col_scip <- method.colour[levels(test_df$ko.method)]


test_p_fdr.plot <- ggplot(mapping = aes(x = P, y = fdp), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = scip_fullnames, values = col_scip)+
  scale_color_manual(labels = scip_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")+
  ylab("FDP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test_p_fdr.plot

test_p_pow.plot <- ggplot(mapping = aes(x = P, y = tpp), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = scip_fullnames, values = col_scip)+
  scale_color_manual(labels = scip_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")+
  ylab("TPP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test_p_pow.plot

mb.semi <- ggarrange(test_p_fdr.plot, test_p_pow.plot, nrow = 1, ncol = 2, common.legend = T)


# plus 
test_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test_p_fdr_plus.plot

test_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test_p_pow_plus.plot

# run time
test_p_time.plot <- ggplot(mapping = aes(x = p_var, y = run.time), data = test_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test_p_time.plot
```
### mvr
```{r mvr test, eval=FALSE}

it = 50
set.seed(1)

test.mvr <- foreach(i = rep(1:it,3), px = rep(c(50, 75, ncol(mb.time1.rm)), each = it), 
                           .packages = c("mixOmics", "tidyverse", "knockoff", "CVglasso","rARPACK", "irlba", "kernlab","JuliaCall")) %dopar%{
                             source(paste0(source.path, "data_sim1.R"))
                             julia_library("Knockoffs")
                             res <- real_mvr(mb.time1.rm, p_x = px, p_s = 15, ko.diagnosis = F, seed = i, call.var = "p_x", q = 0.1)
                           }
real_mvr_df <- do.call("rbind",  test.mvr )
saveRDS(real_mvr_df, "./mb_mvr_94.rds")
```

```{r mvr plot}
real_all_mvr_df <- readRDS("./mb_mvr_94.rds")
real_all_mvr_df$P <- factor(real_all_mvr_df$var, levels = c("50","75","94"))
real_all_mvr_df$ko.method[real_all_mvr_df$ko.method=="sdp_JS"] <- "JS"
real_all_mvr_df$ko.method <- factor(real_all_mvr_df$ko.method)
col_scip <- method.colour[levels(real_all_mvr_df$ko.method)]

real_all_mvr_df$ko.method <- factor(real_all_mvr_df$ko.method)

test.mvr_p_fdr.plot <- ggplot(mapping = aes(x = P, y = fdp), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = mvr_fullnames, values = col_scip)+
  scale_color_manual(labels = mvr_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")+
  ylab("FDP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test.mvr_p_fdr.plot

test.mvr_p_pow.plot <- ggplot(mapping = aes(x = P, y = tpp), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  scale_fill_manual(labels = mvr_fullnames, values = col_scip)+
  scale_color_manual(labels = mvr_fullnames, values = col_scip)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")+
  ylab("TPP") + guides(fill=guide_legend(title="Knockoff Generator"), color=guide_legend(title="Knockoff Generator"))
test.mvr_p_pow.plot

mb.semi.mvr <-  ggarrange(test.mvr_p_fdr.plot, test.mvr_p_pow.plot, nrow = 1, ncol = 2, common.legend = T)

ggarrange(mb.semi, mb.semi.mvr, nrow = 2, ncol = 1, common.legend = F)

# plus 
test.mvr_p_fdr_plus.plot <- ggplot(mapping = aes(x = p_var, y = fdp.plus), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour = ko.method), fun.y="mean", geom="line") +
  scale_y_continuous(breaks = c(0,0.05,0.1,0.2,0.4,0.6,0.8,1))+
  theme(legend.position = "top")+
  geom_hline(yintercept=0.1, linetype="dashed", color = "red")
test.mvr_p_fdr_plus.plot

test.mvr_p_pow_plus.plot <- ggplot(mapping = aes(x = p_var, y = tpp.plus), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr_p_pow_plus.plot

# run time
test.mvr_p_time.plot <- ggplot(mapping = aes(x = p_var, y = run.time), data = real_all_mvr_df)+
  geom_boxplot(aes(fill = ko.method))+
  # scale_fill_manual(values = bench.col)+
  # scale_color_manual(values = bench.col)+
  stat_summary(aes(group= ko.method, colour= ko.method), fun.y="mean", geom="line") +
  theme(legend.position = "top")
test.mvr_p_time.plot
```


# real data analysis
## Transcriptomcs
```{r data process}
multiomics_cpm <- read_csv("./data/multiomics_cpm_matched.csv")

# filter low expression: at least 0.5 cpm in at least 75% of samples
keep.gene <- apply(multiomics_cpm[,c(3:47,49:113)], 1, function(x){sum(x>0.5)>0.75*110})

multiomics_filter <- multiomics_cpm[keep.gene,]
rownames(multiomics_filter) <- multiomics_filter$gene_num


# how many found elevated in placenta
placenta_elevated <- read_tsv("../cfRNA_semi_casestudy/data/tissue_category_rna_placenta_elevated.tsv")

multiomics_cpm_elevated <- multiomics_filter[multiomics_filter$gene_num %in% placenta_elevated$Ensembl,]
#77 genes were found

# kept time 1 sample only
RNA_sample_sub <- substr(colnames(multiomics_cpm)[3:113], 9, 13)
RNA_time1 <- RNA_sample_sub[!duplicated(RNA_sample_sub)] 
#all(RNA_time1 %in% sample.time1$Subject_ID)

cpm_elevated_time1 <- multiomics_cpm_elevated[,c(F,F,!duplicated(RNA_sample_sub))]
RNA_time1_sample <- sample.time1[sample.time1$Subject_ID %in% RNA_time1,]
#all(RNA_time1_sample$Subject_ID == RNA_time1)
colnames(cpm_elevated_time1) <- RNA_time1
rownames(cpm_elevated_time1) <- rownames(multiomics_filter)[multiomics_filter$gene_num %in% placenta_elevated$Ensembl]

```

```{r plsko cpm}

cl <- makeCluster(11) 
registerDoParallel(cl)

it = 50
set.seed(1)

plot(pca(t(cpm_elevated_time1), ncomp =10)$cum.var)
r_emp <- r_criterion(t(cpm_elevated_time1), rmax = 6)

RNA.plsko.placenta <- foreach(i = 1:it, .inorder = T, .errorhandling = "stop", .packages = c("glmnet","knockoff","MASS","rARPACK","mixOmics","CVglasso", 
                    "glmnet","irlba","kernlab")) %dopar% {
                      
                      ko <- knockoffX.sim(t(cpm_elevated_time1), method = "PLSKO.only", threshold.q = 0, r = r_emp)
                      S <- ko.filter(X = t(cpm_elevated_time1), Xk = ko$ko, y = RNA_time1_sample$SampleGroup, q = 0.05, method = "lasso.logistic")
                      return(list(ko = ko, S = S))
                    }

saveRDS(RNA.plsko.placenta, "RNA_plsko_placenta.rds")

RNA.plsko.placenta <- readRDS("RNA_plsko_placenta.rds")
s.placenta <- lapply(RNA.plsko.placenta , function(x){x$S$S})
s.placenta.freq <- as.data.frame(table(unlist(s.placenta))) %>% mutate(gene_num = rownames(cpm_elevated_time1)[as.numeric(as.character(Var1))])

```

```{r cpm ako}
cpm.ko.list <- lapply(RNA.plsko.placenta, function(x){
  x$ko$ko
})

cpm.ako <- AKO.wko(X = t(cpm_elevated_time1), y = RNA_time1_sample$SampleGroup, Xko.list = cpm.ko.list, w.method = "lasso.logistic")
saveRDS(cpm.ako, "cpm_ako.rds")
cpm.ako <- readRDS("cpm_ako.rds")
colnames(t(cpm_elevated_time1))[cpm.ako$ako.s]

```

### transcriptomics HPA v23
```{r cpm v23}
placenta_elevated_v23 <- readRDS("C:/Users/guannany/OneDrive - The University of Melbourne/PhD/database/RNA type/placenta_3fold.rds")

multiomics_cpm_elevated_v23 <- multiomics_filter[multiomics_filter$gene_num %in% placenta_elevated_v23$Gene,]
#86 genes were found

# kept time 1 sample only
RNA_sample_sub <- substr(colnames(multiomics_cpm)[3:113], 9, 13)
RNA_time1 <- RNA_sample_sub[!duplicated(RNA_sample_sub)] 
#all(RNA_time1 %in% sample.time1$Subject_ID)

cpm_elevated_v23_time1 <- multiomics_cpm_elevated_v23[,c(F,F,!duplicated(RNA_sample_sub))]
RNA_time1_sample <- sample.time1[sample.time1$Subject_ID %in% RNA_time1,]
#all(RNA_time1_sample$Subject_ID == RNA_time1)
colnames(cpm_elevated_v23_time1) <- RNA_time1
rownames(cpm_elevated_v23_time1) <- rownames(multiomics_filter)[multiomics_filter$gene_num %in% placenta_elevated_v23$Gene]
```


```{r cmp v23 plsko}
cl <- makeCluster(11) 
registerDoParallel(cl)

plot(pca(t(cpm_elevated_v23_time1), ncomp =10)$cum.var)
r_emp <- r_criterion(t(cpm_elevated_v23_time1), rmax = 5)

cpm_v23.ako <- AKO(X = t(cpm_elevated_v23_time1), n_bootstraps = 50, ko.method = "PLSKO.only",
                        y = RNA_time1_sample$SampleGroup,
                        w.method = "lasso.logistic", threshold.q = 0, r = r_emp)
saveRDS(cpm_v23.ako, "cpm_v23_ako.rds")

cpm_v23.ako <- readRDS("cpm_v23_ako.rds")
colnames(t(cpm_elevated_v23_time1))[cpm_v23.ako$ako.s]

table(unlist(cpm_v23.ako$s))
```


#### limma clipper cpm v23
```{r limma cpm v23}
cpm.dge <- DGEList(cpm_elevated_v23_time1, sample = RNA_time1_sample)
design <- model.matrix(~cpm.dge$samples$SampleGroup)

v <- voom(cpm.dge, design)
limma.fit <- lmFit(v, design)
limma.efit <- eBayes(limma.fit)
limma.t <- limma.efit$t
summary(decideTests(limma.efit))

limma.t <- cbind(limma.t, decideTests(limma.efit)) 
limma.sig <- limma.t[limma.t[,4]!=0,, drop = F]

row.names(limma.sig)
```

```{r clipper cpm v23}

library(Clipper)
cpm.placenta.pe <- cpm.dge$counts[,cpm.dge$samples$SampleGroup=="PE"]
cpm.placenta.control <- cpm.dge$counts[,cpm.dge$samples$SampleGroup=="Term"]

cpm_clipper.selected <- lapply(1:50, function(x){Clipper(cpm.placenta.pe, cpm.placenta.control, analysis = "d", procedure = "GZ", n.permutation = 2, seed = x, contrast.score = "diff")$discoveries[[1]]})
saveRDS(cpm_clipper.selected, "cpm_v23_placenta_clipper.rds")

placenta_clipper.freq <- as.data.frame(table(unlist(cpm_clipper.selected)))
```

```{r wilcoxon cpm v23}
wilcox.placenta.cpm <- lapply(1:nrow(cpm_elevated_v23_time1), function(gene.i){
  wilcox.test(cpm.placenta.control[gene.i,], cpm.placenta.pe[gene.i,])$p.value
})

wilcox.placenta.p.adj <- p.adjust(unlist(wilcox.placenta.cpm))
which(wilcox.placenta.p.adj<0.05)

rownames(cpm_elevated_time1$counts)[which(wilcox.placenta.p.adj<0.05)]

```

## proteins
```{r prot plsko}
prot_elevated <- prot.time1.batrm[,colnames(prot.time1.batrm) %in% placenta_elevated$Gene]

cl <- makeCluster(11) 
registerDoParallel(cl)

it = 50
set.seed(1)

plot(pca(prot_elevated, ncomp =10)$cum.var)
r_emp <- r_criterion(t(log_cpm_elevated_time1), rmax = 5)

prot.plsko.placenta <- foreach(i = 1:it, .inorder = T, .errorhandling = "stop", .packages = c("glmnet","knockoff","MASS","rARPACK","mixOmics","CVglasso", 
                    "glmnet","irlba","kernlab")) %dopar% {
                      
                      ko <- knockoffX.sim(prot_elevated, method = "PLSKO.only", threshold.q = 0, r = r_emp)
                      S <- ko.filter(X = prot_elevated, Xk = ko$ko, y = proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup, q = 0.05, method = "lasso.logistic")
                      return(list(ko = ko, S = S))
                    }

saveRDS(prot.plsko.placenta, "prot_plsko_placenta.rds")

prot.plsko.placenta <- readRDS("prot_plsko_placenta.rds")
s.prot.placenta <- lapply(prot.plsko.placenta , function(x){x$S$S})
s.prot.placenta.freq <- as.data.frame(table(unlist(s.prot.placenta))) %>% mutate(gene_num = colnames(prot_elevated)[as.numeric(as.character(Var1))])
s.prot.placenta_final <- s.prot.placenta.freq %>% filter(Freq>25)

```

```{r proteome ako}

prot.ko.list <- lapply(prot.plsko.placenta, function(x){
  x$ko$ko
})

prot.ako <- AKO.wko(X = prot_elevated, y = proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup, Xko.list = prot.ko.list, w.method = "lasso.logistic")
saveRDS(prot.ako, "prot_ako.rds")

prot.ako <- readRDS("prot_ako.rds")
colnames(prot_elevated)[prot.ako$ako.s]
table(unlist(prot.ako$s))
```
### protein HPA v23
```{r protein v23 ako}
prot_elevated_23 <- prot.time1.batrm[,colnames(prot.time1.batrm) %in% placenta_elevated_v23$`Gene name`]

cl <- makeCluster(11) 
registerDoParallel(cl)

it = 50
set.seed(1)

plot(pca(prot_elevated_23, ncomp =10)$cum.var)
r_emp <- r_criterion(prot_elevated_23, rmax = 8)

prot_v23.ako <- AKO(X = as.matrix(prot_elevated_23), n_bootstraps = 50, ko.method = "PLSKO.only",
                        y = proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup,
                        w.method = "lasso.logistic", threshold.q = 0, r = r_emp)
saveRDS(prot_v23.ako, "prot_ako_v23.rds")

prot_v23.ako <- readRDS("prot_ako_v23.rds")
colnames(prot_elevated_23)[prot_v23.ako$ako.s]

table(unlist(prot_v23.ako$s))
```

### limma protein v23
```{r limma protein v23}
design <- model.matrix(~proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup)

#v <- voom(cpm.dge, design)
limma.fit <- lmFit(t(prot_elevated_23), design)
limma.efit <- eBayes(limma.fit)
limma.t <- limma.efit$t
summary(decideTests(limma.efit))

limma.t <- cbind(limma.t, decideTests(limma.efit)) 
limma.sig <- limma.t[limma.t[,4]!=0,, drop = F]

row.names(limma.sig)
```

```{r protein clipper v23}
library(Clipper)
prot.placenta.pe <- t(prot_elevated_23)[,proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup=="PE"]
prot.placenta.control <- t(prot_elevated_23)[,proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup=="Term"]

prot_clipper.selected <- lapply(1:50, function(x){Clipper(prot.placenta.pe, prot.placenta.control, analysis = "d", procedure = "GZ", n.permutation = 2, seed = x, contrast.score = "diff")$discoveries[[1]]})
saveRDS(prot_clipper.selected, "prot_placenta_clipper_v23.rds")

prot_clipper.selected <- readRDS("prot_placenta_clipper_v23.rds")
prot_clipper.freq <- as.data.frame(table(unlist(prot_clipper.selected)))
```

```{r wilcoxon prot v23}
wilcox.placenta.prot <- lapply(1:ncol(prot_elevated_23), function(gene.i){
  wilcox.test(prot.placenta.pe[gene.i,], prot.placenta.control[gene.i,])$p.value
})

wilcox.placenta.p.adj <- p.adjust(unlist(wilcox.placenta.prot))
which(wilcox.placenta.p.adj<0.05)

colnames(prot_elevated_23)[which(wilcox.placenta.p.adj<0.05)]

```

### protein with placenta release
```{r prot placenta release}
prot_placenta_rel <-  readxl::read_excel("data/12916_2022_2415_MOESM2_ESM.xlsx", 
    sheet = "Placenta-specific released prot")

prot_placenta_df <- prot.time1[,which(colnames(prot.time1) %in% prot_placenta_rel$GeneSymbol)]
```

```{r prot release plsko}
cl <- makeCluster(11) 
registerDoParallel(cl)

plot(pca(prot_placenta_df, ncomp =10)$cum.var)
r_emp <- r_criterion(t(prot_placenta_df), rmax = 10)

prot_pla_rel.ako <- AKO(X = as.matrix(prot_placenta_df), n_bootstraps = 50, ko.method = "PLSKO.only",
                        y = proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup,
                        w.method = "lasso.logistic", threshold.q = 0, r = r_emp)
saveRDS(prot_pla_rel.ako, "prot_ako_placenta_release.rds")

prot_pla_rel.ako <- readRDS("prot_ako_placenta_release.rds")
colnames(prot_placenta_df)[prot_pla_rel.ako$ako.s]

table(unlist(prot_pla_rel.ako$s))
```

```{r protein mvr}
prot_pla_rel.ako_mvr <- AKO(X = as.matrix(prot_placenta_df), n_bootstraps = 50, mvr=T,
                        y = proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup,
                        w.method = "lasso.logistic")
saveRDS(prot_pla_rel.ako_mvr, "prot_ako_placenta_release_mvr.rds")

colnames(prot_placenta_df)[prot_pla_rel.ako_mvr$ako.s]

table(unlist(prot_pla_rel.ako_mvr$s))
```

## microbiome
```{r mb plsko}
mb.group <- as.factor(ifelse(microbiome_data_raw$Preeclampsia[microbiome_data_raw$timepoint==1], "PE", "control"))

cl <- makeCluster(11) 
registerDoParallel(cl)

it = 50
set.seed(1)

plot(pca(mb.time1.rm, ncomp =10)$cum.var)
r_emp <- r_criterion(mb.time1.rm, rmax = 8)

mb.plsko <- foreach(i = 1:it, .inorder = T, .errorhandling = "stop", .packages = c("glmnet","knockoff","MASS","rARPACK","mixOmics","CVglasso", 
                    "glmnet","irlba","kernlab")) %dopar% {
                      
                      ko <- knockoffX.sim(mb.time1.rm, method = "PLSKO.only", threshold.q = 0, r = r_emp)
                      S <- ko.filter(X = mb.time1.rm, Xk = ko$ko, y = mb.group, q = 0.05, method = "lasso.logistic")
                      return(list(ko = ko, S = S))
                    }

saveRDS(mb.plsko, "mb_plsko.rds")

mb.plsko <- readRDS("mb_plsko.rds")
s.mb <- lapply(mb.plsko , function(x){x$S$S})
s.mb.freq <- as.data.frame(table(unlist(s.mb))) %>% mutate(gene_num = colnames(mb.time1.batrm)[as.numeric(as.character(Var1))])
s.mb.freq.final <- s.mb.freq %>% filter(Freq>25)

```

```{r microbio ako}
mb.group <- as.factor(ifelse(microbiome_data_raw$Preeclampsia[microbiome_data_raw$timepoint==1], "PE", "control"))

cl <- makeCluster(11) 
registerDoParallel(cl)

plot(pca(mb.time1.rm, ncomp =10)$cum.var)
r_emp <- r_criterion(mb.time1.rm, rmax = 10)

mb.ako <- AKO(X = mb.time1.rm, n_bootstraps = 50, ko.method = "PLSKO.only",
                        y = mb.group,
                        w.method = "lasso.logistic", threshold.q = 0, r = r_emp)
saveRDS(mb.ako, "mb_ako.rds")

mb.ako <- readRDS("mb_ako.rds")
# s.mb.ako.taxa <- assignTaxonomy(colnames(mb.time1.batrm)[mb.ako$ako.s], refFasta = "silva_nr99_v138.1_wSpecies_train_set.fa")

table(unlist(mb.ako$s))

mb.taxa <- readRDS("microbiome_taxa_final.rds")
s.mb.ako.taxa <- mb.taxa[c(67,38, 83, 1),]
```


# limma and clipper

## transcript
```{r limma}
cpm.dge <- DGEList(cpm_elevated_time1, sample = RNA_time1_sample)
design <- model.matrix(~cpm.dge$samples$SampleGroup)

v <- voom(cpm.dge, design)
limma.fit <- lmFit(v, design)
limma.efit <- eBayes(limma.fit)
limma.t <- limma.efit$t
summary(decideTests(limma.efit))

limma.t <- cbind(limma.t, decideTests(limma.efit)) 
limma.sig <- limma.t[limma.t[,4]!=0,, drop = F]

row.names(limma.sig)
```

```{r clipper cpm}

library(Clipper)
cpm.placenta.pe <- cpm.dge$counts[,cpm.dge$samples$SampleGroup=="PE"]
cpm.placenta.control <- cpm.dge$counts[,cpm.dge$samples$SampleGroup=="Term"]

cpm_clipper.selected <- lapply(1:50, function(x){Clipper(cpm.placenta.pe, cpm.placenta.control, analysis = "d", procedure = "GZ", n.permutation = 2, seed = x, contrast.score = "diff")$discoveries[[1]]})
saveRDS(cpm_clipper.selected, "cpm_placenta_clipper.rds")

placenta_clipper.freq <- as.data.frame(table(unlist(cpm_clipper.selected)))
```

```{r wilcoxon cpm}
wilcox.placenta.cpm <- lapply(1:77, function(gene.i){
  wilcox.test(cpm.placenta.control[gene.i,], cpm.placenta.pe[gene.i,])$p.value
})

wilcox.placenta.p.adj <- p.adjust(unlist(wilcox.placenta.cpm))
which(wilcox.placenta.p.adj<0.05)

rownames(cpm_elevated_time1$counts)[which(wilcox.placenta.p.adj<0.05)]

```

## proteome
```{r limma protein}
design <- model.matrix(~proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup)

#v <- voom(cpm.dge, design)
limma.fit <- lmFit(t(prot_elevated), design)
limma.efit <- eBayes(limma.fit)
limma.t <- limma.efit$t
summary(decideTests(limma.efit))

limma.t <- cbind(limma.t, decideTests(limma.efit)) 
limma.sig <- limma.t[limma.t[,4]!=0,, drop = F]

row.names(limma.sig)
```

```{r protein clipper}
library(Clipper)
prot.placenta.pe <- t(prot_elevated)[,proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup=="PE"]
prot.placenta.control <- t(prot_elevated)[,proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup=="Term"]

prot_clipper.selected <- lapply(1:50, function(x){Clipper(prot.placenta.pe, prot.placenta.control, analysis = "d", procedure = "GZ", n.permutation = 2, seed = x, contrast.score = "diff")$discoveries[[1]]})
saveRDS(prot_clipper.selected, "prot_placenta_clipper.rds")

prot_clipper.selected <- readRDS("prot_placenta_clipper.rds")
prot_clipper.freq <- as.data.frame(table(unlist(prot_clipper.selected)))
```

```{r wilcoxon prot}
wilcox.placenta.prot <- lapply(1:63, function(gene.i){
  wilcox.test(prot.placenta.pe[gene.i,], prot.placenta.control[gene.i,])$p.value
})

wilcox.placenta.p.adj <- p.adjust(unlist(wilcox.placenta.prot))
which(wilcox.placenta.p.adj<0.05)

colnames(prot_elevated)[which(wilcox.placenta.p.adj<0.05)]

```

```{r limma prot release}
design <- model.matrix(~proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup)

#v <- voom(cpm.dge, design)
limma.fit <- lmFit(t(prot_placenta_df), design)
limma.efit <- eBayes(limma.fit)
limma.t <- limma.efit$t
summary(decideTests(limma.efit))

limma.t <- cbind(limma.t, decideTests(limma.efit)) 
limma.sig <- limma.t[limma.t[,4]!=0,, drop = F]

row.names(limma.sig)
```

```{r protein clipper release}
library(Clipper)
prot.placenta.pe <- t(prot_placenta_df)[,proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup=="PE"]
prot.placenta.control <- t(prot_placenta_df)[,proteome_data_raw[proteome_data_raw$TimePoint=="1st Tri",]$SampleGroup=="Term"]

prot_clipper.selected <- lapply(1:50, function(x){Clipper(prot.placenta.pe, prot.placenta.control, analysis = "d", procedure = "GZ", n.permutation = 2, seed = x, contrast.score = "diff")$discoveries[[1]]})
saveRDS(prot_clipper.selected, "prot_release_clipper.rds")

prot_clipper.selected <- readRDS("prot_release_clipper.rds")
prot_clipper.freq <- as.data.frame(table(unlist(prot_clipper.selected)))
```

```{r wilcoxon prot release}
wilcox.release.prot <- lapply(1:36, function(gene.i){
  wilcox.test(prot.placenta.pe[gene.i,], prot.placenta.control[gene.i,])$p.value
})

wilcox.placenta.p.adj <- p.adjust(unlist(wilcox.release.prot))
which(wilcox.placenta.p.adj<0.05)

colnames(prot_placenta_df)[which(wilcox.placenta.p.adj<0.05)]

```

## microbio
```{r limma mb}
# TMM normalisation
mb.tmm <- t(microbiome.keep) %>% DGEList(samples = microbiome_data_raw$Preeclampsia) %>% calcNormFactors(method = "TMM")
class(mb.clr) <- 'matrix' 

# filter time 1 samples
mb.time1.tmm <- mb.tmm[,microbiome_data_raw$timepoint==1]
mb.time1.tmm <- mb.time1.tmm$counts[rownames(mb.time1.tmm$counts) %in% colnames(mb.time1.rm),]

mb.tmm.dge <- DGEList(mb.time1.tmm, sample = microbiome_data_raw$Preeclampsia[microbiome_data_raw$timepoint==1])

design <- model.matrix(~mb.tmm.dge$samples$samples)

v <- voom(mb.tmm.dge, design)
limma.fit <- lmFit(v, design)
limma.efit <- eBayes(limma.fit)
limma.t <- limma.efit$t
summary(decideTests(limma.efit))

limma.t <- cbind(limma.t, decideTests(limma.efit)) 
limma.sig <- limma.t[limma.t[,4]!=0,, drop = F]

row.names(limma.sig)
```

```{r wilcoxon mb}
mb.time1.rm.pe <- mb.time1.rm[mb.group=="PE",] 
mb.time1.rm.control <- mb.time1.rm[mb.group =="control",]

wilcox.mb.p <- lapply(1:ncol(mb.time1.rm), function(gene.i){
  wilcox.test(mb.time1.rm.control[,gene.i], mb.time1.rm.pe[,gene.i])$p.value
})

wilcox.mb.p.adj <- p.adjust(unlist(wilcox.mb.p))
which(wilcox.mb.p.adj<0.05)

colnames(mb.time1.batrm)[which(wilcox.mb.p.adj<0.05)]
```
```{r clipper mb}
library(Clipper)

mb_clipper.selected <- lapply(1:50, function(x){Clipper(t(mb.time1.batrm.pe), t(mb.time1.batrm.control), analysis = "d", procedure = "GZ", n.permutation = 2, seed = x, contrast.score = "diff")$discoveries[[1]]})

mb.freq.clipper <- as.data.frame(table(unlist(mb_clipper.selected)))
mb.freq.clipper.final<- mb.freq.clipper %>% filter(Freq > 25) %>% mutate(name =colnames(mb.time1.batrm.pe)[as.numeric(as.character(Var1))]) %>% arrange(desc(Freq))
```

```{r mb tmm plsko}

mb.group <- as.factor(ifelse(microbiome_data_raw$Preeclampsia[microbiome_data_raw$timepoint==1], "PE", "control"))
mb.tmm.log <- t(microbiome.keep) %>% DGEList(samples = microbiome_data_raw$Preeclampsia) %>% calcNormFactors(method = "TMM", log = T)
mb.tmm.log.time1 <- mb.tmm.log[,microbiome_data_raw$timepoint==1]
mb.time1.tmm.log <- mb.tmm.log.time1$counts[rownames(mb.tmm.log.time1$counts) %in% colnames(mb.time1.rm),]


cl <- makeCluster(11) 
registerDoParallel(cl)

plot(pca(t(mb.tmm.log.time1$counts), ncomp =10)$cum.var)
r_emp <- r_criterion(t(mb.tmm.log.time1$counts), rmax = 4)

mb.tmm.ako <- AKO(X = t(mb.tmm.dge$counts), n_bootstraps = 50, ko.method = "PLSKO.only",
                        y = mb.group,
                        w.method = "lasso.logistic", threshold.q = 0, r = r_emp)
saveRDS(mb.tmm.ako, "mb_tmm_ako.rds")

mb.tmm.ako <- readRDS("mb_tmm_ako.rds")
# s.mb.ako.taxa <- assignTaxonomy(colnames(mb.time1.batrm)[mb.ako$ako.s], refFasta = "silva_nr99_v138.1_wSpecies_train_set.fa")

table(unlist(mb.tmm.ako$s))

mb.taxa <- readRDS("microbiome_taxa_final.rds")
s.mb.tmm.ako.taxa <- mb.taxa[mb.tmm.ako$ako.s,]
s.mb.tmm.ako.taxa <- mb.taxa[c(1,2),]

```

